### 2015-04-08 approach.tex
Bringing everything up to date

A high level description of our approach is outlined in \FG{fig:pipeline}. From the oriented surface samples, we construct the approximation $\vec{V}(\vx)$ using our variational formulation. We then apply FIR filters to the coefficients of the approximation to construct a scalar field representing the approximate divergence of the smoothed normal field (\FG{fig:pipeline}c.) We then use an appropriate discretization of the inverse Laplacian operator to solve the Poisson equation \EQ{eq:poisson}, and obtain a smooth function that approximates the original indicator function (\FG{fig:pipeline}c). Our code will is also available for download

\subsection{Obtaining $\vec{V}(\vx)$} \label{sec:getting_vd}
\label{sec:obtainingV}
We define the vector field as $\vec{V}(\vx):=\mathbf{L} \left[\upsilon_1(\vx) \ \upsilon_2(\vx) \ \upsilon_3(\vx)\right]^T$ where $\mathbf{L := [b_1 \ b_2 \ b_3]}$ is a basis for $\realn^3$ that contains the principal lattice directions of the chosen lattice. Each $\upsilon_i$ is a scalar function representing the projection of the smoothed normal field onto each respective principal lattice direction $\mathbf{b}_i$. This may seem unintuitive at first, however, using a non-canonical basis to represent $\vec{V}$ will be convenient in two ways. Firstly, it will allow us to approximate the divergence of $\vec{V}$ by taking directional derivatives of $\vec{V}$ in the principal lattice directions. This is motivated by the fact that it is more natural to approximate derivatives in the principal lattice directions on non-Cartesian lattices. Secondly, it will allow us to construct divergence of the smoothed normal field from more than three principal lattice directions. We also allow each function $\upsilon_i$ to potentially be represented in different function spaces, i.e. {\small 
\begin{equation} 
	\upsilon_i(\vx) = \sum_{k=0}^N \text{v}_k^i \varphi_k^i(\vx) \in \fspce{\varphi^i}{h\mathbf{L}}.
\end{equation}}
The coefficient vector of $\upsilon_i$ is denoted as $\mathbf{v}_i:=\left[\text{v}_1^i \ \cdots \ \text{v}_N^i \right]$. We consider each component of $\vec{V}(\vx)$ separately, thus, it suffices to only look at one $\upsilon_i$. We simply find the $\upsilon_i$ that minimizes {\small 
\begin{equation} \label{eq:vari_eqn}
 	\sump_{\forall (\vx,\vn) \in P }^{M} (\upsilon_i(\vx) - (\mathbf{L}^{-1})_i \cdot \mathbf{n})^2 + E(\upsilon_i).
\end{equation} }
Where $(\mathbf{L}^{-1})_i$ is the $i^{th}$ column of $(\mathbf{L}^{-1}).$ As before, this provides us with a parametrized way of controlling the smoothness of the resulting approximation, without the need to explicitly choose a smoothing kernel.

%In what follows, $\mathbf{L}$ is chosen to be the canonical basis $\mathbf{I_3}$ for the Cartesian lattice, whereas for the BCC lattice, we choose $\mathbf{L = L}_B$.

\subsection{Divergence of $\vec{V}(\vx)$}
\label{sec:divergence}
A simple way to obtain the divergence of $\vec{V}(\vx)$ is to consider its analytical partial derivatives which, due to the linearity of the expression, requires one to only take the analytic partial derivatives of the underlying kernel functions. However, this reduces the smoothness of the solution \cite{zahgrad}. There is also no conclusive evidence that suggests the analytic derivative of the kernel provides a better approximation to true underlying derivative. With this in mind, we choose to take a discrete filtering approach and approximate the divergence of $\vec{V}$ with a set of discrete FIR filters. While these filters do not strictly facilitate the best approximation scheme for derivatives, they provide a good compromise between compactness versus accuracy, even more so when the reconstruction spaces are shifted \cite{gradrev}.

\subsubsection{Spaces Spanned by a Single Generating Function}
When each space is identical (when $\upsilon_i \in \fspce{\varphi}{L}$) the divergence can be found by taking directional derivatives of $\vec{V}$ along the vectors in $\mathbf{L}$ as follows: {\small 
\begin{eqnarray}\label{eq:div_dir}
	\grad \cdot \vec{V}(\vx) & = & \grad \cdot \mathbf{L} \left[\upsilon_1(\vx) \ \upsilon_2(\vx) \ \upsilon_3(\vx)\right]^T \notag \\
	 & = & \left[ \mathbf{(\grad \cdot b_1) \ (\grad \cdot b_2) \ (\grad \cdot b_3 ) }\right] \cdot \left[\upsilon_1 \ \upsilon_2 \ \upsilon_3\right]^T \\
	 & = & \dv{\mathbf{b_1}}\upsilon_1 + \dv{\mathbf{b_2}}\upsilon_2 + \dv{\mathbf{b_3}}\upsilon_3. \notag
\end{eqnarray} }
This allows us to approximate the divergence of $\vec{V}$ by simply convolving the coefficient vector corresponding to $\upsilon_i$ with an appropriate 1D derivative filter $d[n]$, then sum the coefficients. The expression for the resulting coefficient vector of the divergence is then {\small
\begin{equation}
	\mathbf{f}:= \sum_{i=1}^3{ \mathbf{v}_i*\mathbf{d}_i}
\label{eq:divergenceViaFiltering}
\end{equation}}
where $\mathbf{d}_i$ is a directional derivative filter obtained by orienting the 1D filter $d$ 
along the direction $\mathbf{b}_i.$

>>>>

A high level description of our approach is outlined in \FG{fig:pipeline}. From the oriented surface samples, we construct the approximation $\vec{V}(\vx)$ using our variational formulation. We then apply FIR filters to the coefficients of the approximation to construct a scalar field representing the approximate divergence of the smoothed normal field (\FG{fig:pipeline}c.) We then use an appropriate discretization of the inverse Laplacian operator to solve the Poisson equation \EQ{eq:poisson}, and obtain a smooth function that approximates the original indicator function (\FG{fig:pipeline}c). Our code is also freely available for download.

\subsection{Obtaining $\vec{V}(\vx)$} \label{sec:getting_vd}
\label{sec:obtainingV}
We define the vector field as $\vec{V}(\vx):=\mathbf{L} \left[\upsilon_1(\vx) \ \upsilon_2(\vx) \ \upsilon_3(\vx)\right]^T$ where $\mathbf{L := [b_1 \ b_2 \ b_3]}$ is a basis for $\realn^3$ that contains the principal lattice directions of the chosen lattice. Each $\upsilon_i$ is a scalar function representing the projection of the smoothed normal field onto each respective principal lattice direction $\mathbf{b}_i$. This may seem unintuitive at first, however, using a non-canonical basis to represent $\vec{V}$ will be convenient in two ways. Firstly, it will allow us to approximate the divergence of $\vec{V}$ by taking directional derivatives of $\vec{V}$ in the principal lattice directions. This is motivated by the fact that it is more natural to approximate derivatives in the principal lattice directions on non-Cartesian lattices. Secondly, it will allow us to construct divergence of the smoothed normal field from more than three principal lattice directions. We also allow each function $\upsilon_i$ to potentially be represented in different function spaces, i.e. {\small 
\begin{equation} 
	\upsilon_i(\vx) = \sum_{k=0}^N \text{v}_ki \varphi_k^i(\vx) \in \aspace{\lattice{L}{\varphi^i}}.
\end{equation}}
The coefficient vector of $\upsilon_i$ is denoted as $\mathbf{v}_i:=\left[\text{v}_1^i \ \cdots \ \text{v}_N^i \right]$.We use the unfortunate notation $\text{v}_j^i, \varphi_k^i$ to denote that those objects belong to $\upsilon_i(x)$, not to denote exponentiation. We consider each component of $\vec{V}(\vx)$ separately, thus, it suffices to describe the process at one $\upsilon_i$. We simply find the $\upsilon_i$ that minimizes {\small 
\begin{equation} \label{eq:vari_eqn}
 	\sump_{\forall (\vx,\vn) \in P }^{M} (\upsilon_i(\vx) - (\mathbf{L}^{-1})_i \cdot \mathbf{n})^2 + E(\upsilon_i).
\end{equation} }
Where $(\mathbf{L}^{-1})_i$ is the $i^{th}$ column of $(\mathbf{L}^{-1}).$ As before, this provides us with a parametrized way of controlling the smoothness of the resulting approximation, without the need to explicitly choose a smoothing kernel. These $\mathbf{v}_i$ can be obtained using the procedure of \SC{sec:vari_review}.

\subsection{Divergence of $\vec{V}(\vx)$}
\label{sec:divergence}
A simple way to obtain the divergence of $\vec{V}(\vx)$ is to consider its analytical partial derivatives which, due to the linearity of the expression, requires one to only take the analytic partial derivatives of the underlying kernel functions. However, this reduces the smoothness of the solution \cite{zahgrad}, and there is also no conclusive evidence that suggests the analytic derivative of the kernel provides a better approximation to true underlying derivative. With this in mind, we choose to take a discrete filtering approach and approximate the divergence of $\vec{V}$ with a set of discrete FIR filters. While these filters do not strictly facilitate the best approximation scheme for derivatives, they provide a good compromise between compactness versus accuracy, even more so when the reconstruction spaces are shifted \cite{gradrev}.

\subsubsection{Spaces Spanned by a Single Generating Function}
When each space is identical (when $\upsilon_i \in \aspace{\lattice{L}_h}{\varphi}$) the divergence can be found by taking directional derivatives of $\vec{V}$ along the vectors in $\mathbf{L}$ as follows: {\small 
\begin{eqnarray}\label{eq:div_dir}
	\grad \cdot \vec{V}(\vx) & = & \grad \cdot \mathbf{L} \left[\upsilon_1(\vx) \ \upsilon_2(\vx) \ \upsilon_3(\vx)\right]^T \notag \\
	 & = & \left[ \mathbf{(\grad \cdot b_1) \ (\grad \cdot b_2) \ (\grad \cdot b_3 ) }\right] \cdot \left[\upsilon_1 \ \upsilon_2 \ \upsilon_3\right]^T \\
	 & = & \dv{\mathbf{b_1}}\upsilon_1 + \dv{\mathbf{b_2}}\upsilon_2 + \dv{\mathbf{b_3}}\upsilon_3. \notag
\end{eqnarray} }
This allows us to approximate the divergence of $\vec{V}$ by simply convolving the coefficient vector corresponding to $\upsilon_i$ with an appropriate 1D derivative filter $d[n]$, then sum the coefficients. The expression for the resulting coefficient vector of the divergence is then {\small
\begin{equation}
	\mathbf{f}:= \sum_{i=1}^3{ \mathbf{v}_i*\mathbf{d}_i}
\label{eq:divergenceViaFiltering}
\end{equation}}
where $\mathbf{d}_i$ is a directional derivative filter obtained by orienting the 1D filter $d$ 
along the direction $\mathbf{b}_i.$

### 2015-04-08 indicator.tex
Bringing everything up to date

\section{Recovering the Indicator}
\label{sec:recoveringX}
Once the divergence of the smoothed vector field has been computed, the remaining step is to find a function that satisfies equation \EQ{eq:poisson}. As with the computation of a function's derivative, we also seek a filtering solution for the inverse of the Laplacian. We desire a filter $\mathbf{l}^{-1}[\mathbf{m}]$ that provides a fourth-order discretization of the inverse Laplacian operator $\Delta^{-1}$. This can be achieved by first obtaining a 1D filter $l$ that provides a fourth-order discretization of the 1D second-derivative operator. On the Cartesian and BCC lattices, application of this filter along the principal directions followed by a summation of the resulting coefficients, yields a fourth-order discretization of the Laplacian~\cite{usmanthesis}. The inverse Laplacian filter $\mathbf{l}^{-1}$ is simply the inverse of the resulting Laplacian filter. Our approximation $\inda(\vx)$ can now be written as {\small
\begin{equation}
	\inda(\vx)  = \displaystyle\sum_{\mathbf{k}} (\mathbf{f} \ast \mathbf{l}^{-1})_k\varphi_k(\vx)
\end{equation}}
where $\mathbf{f}$ is the coefficient vector obtained in the previous step (cf. equation~\EQ{eq:divergenceViaFiltering}). The only stipulations to this is that the indicator function must satisfy zero Dirichlet boundary conditions (which is given as part of the problem definition). In practice, the filter $\mathbf{l}^{-1}$ does not need to be computed explicitly as it can be efficiently applied in the Fourier domain via a Discrete Sine Transform (DST) that only depends on the 1D filter $l$~\cite[Chapter 4]{usmanthesis}. Whereas the DST on the Cartesian lattice is well-defined, on the BCC lattice, we utilize a  modified DST that can be efficiently implemented via 1D DSTs~\cite{bccdst}. The particular filter $l$ used in our experiments can be found in \TA{tab:filters}.

\subsection{Extracting the Iso-Contour}\label{sec:levelset}
With a true indicator function, the level-set desired is the set of points $\vx$ such that $\inda(\vx) = \frac{1}{2}$, which can readily be visualized either via marching algorithms, ray tracing, or other volumetric visualization techniques. However, the constant scaling factor of equation \EQ{eq:vdef} offsets the appropriate level set at which to threshold the indicator function. A common approach that has shown to be quite robust in this situation \cite{fftk}, is to average the levelset obtained by evaluating the approximation at the input points $\frac{1}{|P|}\sum_{(\vx,\vn)\in P}{\tilde{\chi}(\vx)}.$ The reasoning for this, is that the iso-value at the input data should be on, or at least near the surface of the model. Once the iso-surface has been determined, the mesh is extracted via marching cubes, with the step density set to half the lattice scaling factor ($frac{h}{2}$) \cite{}

>>>>

\section{Recovering the Indicator}

\label{sec:recoveringX}
Returning to the surface reconstruction approach, once the divergence of the smoothed vector field has been computed, the remaining step is to find a function that satisfies equation \EQ{eq:poisson}. As with the computation of a function's derivative, we also seek a filtering solution for the inverse of the Laplacian. We desire a filter $\mathbf{l}^{-1}[\mathbf{m}]$ that provides a fourth-order discretization of the inverse Laplacian operator $\Delta^{-1}$. This can be achieved by first obtaining a 1D filter $l$ that provides a fourth-order discretization of the 1D second-derivative operator. On the Cartesian and BCC lattices, application of this filter along the principal directions followed by a summation of the resulting coefficients, yields a fourth-order discretization of the Laplacian~\cite{usmanthesis}. The inverse Laplacian filter $\mathbf{l}^{-1}$ is simply the inverse of the resulting Laplacian filter. Our approximation $\inda(\vx)$ can now be written as {\small
\begin{equation}
	\inda(\vx)  = \displaystyle\sum_{\mathbf{k}} (\mathbf{f} \ast \mathbf{l}^{-1})_k\varphi_k(\vx)
\end{equation}}
where $\mathbf{f}$ is the coefficient vector obtained in the previous step (cf. equation~\EQ{eq:divergenceViaFiltering}). The only stipulations to this is that the indicator function must satisfy zero Dirichlet boundary conditions (which is given as part of the problem definition). In practice, the filter $\mathbf{l}^{-1}$ does not need to be computed explicitly as it can be efficiently applied in the Fourier domain via a Discrete Sine Transform (DST) that only depends on the 1D filter $l$~\cite[Chapter 4]{usmanthesis}. Whereas the DST on the Cartesian lattice is well-defined, on the BCC lattice, we utilize a  modified DST that can be efficiently implemented via 1D DSTs~\cite{bccdst}. 
%The particular filter $l$ used in our experiments can be found in \TA{tab:filters}.

\subsection{Extracting the Iso-Contour}\label{sec:levelset}
With a true indicator function, the level-set desired is the set of points $\vx$ such that $\inda(\vx) = \frac{1}{2}$, which can readily be visualized either via marching algorithms, ray tracing, or other volumetric visualization techniques. However, the constant scaling factor of equation \EQ{eq:vdef} offsets the appropriate level set at which to threshold the indicator function. A common approach that has shown to be quite robust in this situation \cite{fftk}, is to average the levelset obtained by evaluating the approximation at the input points ${1}/{|P|}\sum_{(\vx,\vn)\in P}{\tilde{\chi}(\vx)}.$ The reasoning for this, is that the iso-value at the input data should be on, or at least near the surface of the model. Once the iso-surface has been determined, the mesh is extracted via marching cubes, with the step density set to half the lattice scaling factor ($\frac{h}{2}$) % \cite{}
